A method for defining dispersed community territories

### 2024-03-06 R/community-clustering.qmd
final edits before resubmission

  ggplot() +
    geom_spatraster(
      data = hill, 
      fill = hcl.colors(1000, "Grays")[values(hill)],
      maxcell = Inf
    ) +
    geom_spatraster(data = x[["dem"]], alpha = 0.5) +
    scale_fill_hypso_c(
      name = "Elevation (m)",
      palette = "utah_1",
      limits = c(1350, 3850),
      alpha = 0.5,
      na.value = "transparent"
    ) +
    with_outer_glow(
      geom_sf(
        data = x[["sites"]] |> filter(center == 0),
        shape = 16,
        color = "gray25",
        alpha = 0.5,
        size = 0.6
      ),
      colour = "white",
      sigma = 0.5,
      expand = 1
    ) +
    with_outer_glow(
      geom_sf(
        data = x[["sites"]] |> filter(center == 1),
        shape = 17,
        color = "#CD3C00",
        alpha = 0.8,
        size = 1.1
      ),
      colour = "white",
      sigma = 0.5,
      expand = 1
    ) +
    annotation_scale(
      aes(location = "bl", line_col = "white", text_col = "white"),
      height = unit(0.18, "cm"),
      line_width = 0.5
    ) +
    coord_sf(expand = FALSE) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      legend.position = "none",
      plot.margin = margin(l = 2)
    )

@@

ggsave(
  plot = gg,
  filename = fn,
  width = 7.5, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

@@

  missing_cells <- setdiff(cell_ids, cells_in_graph)

  if (length(missing_cells) > 0){

    cli::cli_abort(c(
      "Some grid cells have sites but are not in the graph!",
      "i" = "Cells: {missing_cells}"
    ))

@@

1.  Identify community centers
2.  Join centers to their overlapping neighbors
3.  Remove distant farms
4.  Join farms to their nearest center
5.  Draw smallest concave hull encompassing all farms, centers, and paths

Step 1 is already done and stored in the data. Step 3 is technically done before
2 since it means those sites can be ignored afterwards.

@@

  hell <- function(x, n = 1000){

    slope <- terrain(x, "slope", unit = "radians")
    aspect <- terrain(x, "aspect", unit = "radians")

    hill <- shade(slope, aspect, 30, 45)
    hill <- setValues(hill, scales::rescale(values(hill), to = c(1, n)))
    hill <- round(hill)

    list(
      shade = hill,
      grays = hcl.colors(n, "Grays")[values(hill)]
    )

  h <- hell(x[["dem"]])

@@

# patchwork is great, but not perfect...
bob <- theme(
  legend.box.margin = margin(t = -10),
  legend.key.size = unit(0.5, "cm"),
  legend.margin = margin(),
  legend.justification = "left",
  legend.position = "bottom"
)

@@

ggsave(
  plot = zg + plot_layout(guides = "collect") & bob,
  filename = fn,
  width = 7.5, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

>>>>

  h <- list(
    shade = hill,
    grays = hcl.colors(1000, "Grays")[values(hill)]
  )

  ggplot() +
    geom_spatraster(
      data = h[["shade"]], 
      fill = h[["grays"]],
      maxcell = Inf
    ) +
    geom_spatraster(data = x[["dem"]], alpha = 0.5) +
    scale_fill_distiller(palette = "Greys", guide = "none") +
    ggnewscale::new_scale_fill() +
    scale_fill_hypso_c(
      name = "Elevation (m)",
      palette = "utah_1",
      limits = c(1350, 3850),
      alpha = 0.5,
      na.value = "transparent"
    ) +
    with_outer_glow(
      geom_sf(
        data = x[["sites"]] |> filter(center == 0),
        shape = 16,
        color = "gray25",
        alpha = 0.5,
        size = 0.8
      ),
      colour = "white",
      sigma = 1,
      expand = 1
    ) +
    with_outer_glow(
      geom_sf(
        data = x[["sites"]] |> filter(center == 1),
        shape = 17,
        color = "#CD3C00",
        alpha = 0.9,
        size = 1
      ),
      colour = "white",
      sigma = 1,
      expand = 1
    ) +
    annotation_scale(
      aes(location = "bl", line_col = "white", text_col = "white"),
      height = unit(0.18, "cm"),
      line_width = 0.5
    ) +
    coord_sf(expand = FALSE) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      legend.position = "none",
      plot.margin = margin(l = 2)
    )

@@

ggsave(
  plot = gg,
  filename = fn,
  width = 7, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

@@

@@

1.  Identify community centers
2.  Join farms to their nearest center
3.  Remove distant farms
4.  Join centers to their overlapping neighbors
5.  Draw smallest concave hull encompassing all farms, centers, and paths

Step 1 is already done and stored in the data. Step 2 is technically done after
step 4. It just makes more sense to think of it coming first, like we're
building *up* to the community - simply put, small things connect to big things,
and big things merge into even bigger things.

@@

  h <- list(
    shade = hill,
    grays = hcl.colors(1000, "Grays")[values(hill)]
  )

@@

# patchwork is great, but not perfect...
bob <- theme(
  legend.box.margin = margin(t = -10),
  legend.key.size = unit(0.43, "cm"),
  legend.margin = margin(),
  legend.justification = "left",
  legend.position = "bottom"
)

@@

ggsave(
  plot = zg + plot_layout(guides = "collect") & bob,
  filename = fn,
  width = 7, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

### 2024-03-06 manuscript/manuscript.qmd
final edits before resubmission

Not surprisingly, virtually all of the Mesa Verde NP communities show high
densities. The community that includes Goodman Point and Sand Canyon Pueblo just
north of Ute Mountain also has a relatively high density, as do several of the
communities west of Yellow Jacket and along the Dolores River. In the NRG, the
highest density community is near the center of the region. It juts out from the
Rio Grande at its confluence with the Pojoaque River and includes the San
Ildefonso and Pojoaque pueblos. Another high density community lies in a shallow
north-south canyon at the base of the Sangre de Cristos. Additional high density
communities are observed on the Pajarito Plateau in Bandelier NM. In the CIB
area, the three densest communities can be found in the far eastern edge of the
study area near El Morro NM, with additional dense communities near Ramah just
to the west of El Morro and Manuelito in the northwest.

>>>>

Not surprisingly, virtually all of the Mesa Verde NP communities show high
densities. The community that includes Goodman Point and Sand Canyon Pueblo just
north of Ute Mountain also has a relatively high density, as do several of the
communities west of Yellow Jacket and along the Dolores River. In the NRG, the
highest density community is near the center of the region. It juts out from the
Rio Grande at its confluence with the Pojoaque River and includes the San
Ildefonso and Pojoaque pueblos. Another high density community lies in a shallow
north-south canyon at the base of the Sangre de Cristos. Additional high density
communities are observed on the Pajarito Plateau in Bandelier NM. In the CIB
area, the three densest communities can be found in the far eastern edge of the
study area near El Morro NM, with additional dense communities near Ramah just
to the west of El Morro, southeast of Zuni near the center of Cibola, and south
of Manuelito in the northwest.
